---
title: "CSF ocrevus samples analysis: inspection of macrophage cluster enriched in treatment"
author: "PPA"
date: "`r Sys.Date()`"
output:
  pdf_document: 
    number_sections: yes
    toc: yes
    toc_depth: 4
  html_document:
    df_print: paged
geometry: margin=0.5in
header-includes:
   - \usepackage{subfig}
   - \usepackage{float}
---



```{r setup, include=FALSE}

##########################
## Packages and options ##
##########################

library(gridExtra)
library(knitr)
library(readr)
library(dplyr)
library(tidyr)
library(tibble)
library(stringr)
library(ggplot2)
library(ggExtra)
library(stats)
# library(RColorBrewer)
library(viridis)
library(scales)
library(ggrepel)
library(circlize)
library(RColorBrewer)
library(gplots)
library(cowplot)
library(gtable)
library(data.table)
library(Seurat)
library(hdf5r)
library(ggsci)
library(ggridges)
library(forcats)
library(kableExtra)
library(sp)
# library(FactoMineR)
# library(factoextra)
library(wesanderson)
library(ComplexHeatmap)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.pos = 'H')
knitr::opts_chunk$set(dev = 'png', dpi = 300)

#set root directory to be one level up of the current directory
#knitr::opts_knit$set(root.dir = '.')

#Caching option when tweaking the knitting
knitr::opts_chunk$set(cache = T)

WRITE_FILES <- FALSE


##################
## ggplot theme ##
##################

theme_mrl <- function(x = 1) {
  theme_minimal() +
    theme(
      axis.line = element_line(),
      axis.ticks.x = element_line(),
      axis.ticks.y = element_line(),
      axis.text.x = element_text(size = 12*x,face = "plain", angle = 40, vjust = 0.9, hjust=0.9, colour = "black"),
      axis.text.y = element_text(size = 12*x,face = "plain", colour = "black"),
      axis.title.x = element_text(size = 12*x,face = "bold"),
      axis.title.y = element_text(size = 12*x,face = "bold"),
      strip.background = element_rect(fill="gray20", colour="gray20", linetype="solid"),
                                      strip.text = element_text(size=12*x, colour="white", face="bold"),
                                      legend.title = element_text(size=14*x, face = "bold"),
                                      legend.text = element_text(size=12*x, color="black", face="bold"),
      legend.background = element_rect(fill = "transparent", colour = "transparent"),
      plot.title =  element_text(hjust=0.5, vjust=2, face="bold"),
      plot.subtitle = element_text(hjust=0.5, vjust=3, face="italic"),
      plot.caption = element_text(hjust = 0, face = "italic")
                                      )
}


opts <- options()  # save old options

options(ggplot2.continuous.colour="viridis")
options(ggplot2.continuous.fill = "viridis")


palette <- c(pal_rickandmorty()(12),pal_jco()(10), pal_futurama()(11), pal_tron()(7))

options(bitmapType = 'cairo', device = 'png')
```


There's a suspicion that the macrophage cluster has contamination of T cells in it. Is it really it or:
* myeloid cells with T cell features (and vice versa)
* doublets of T and myeloid
* contamination by ambient RNA

\pagebreak


# Classification of T cell contaminants in the macrophage cluster

\  


```{r}
seurat <- readRDS("./results/CSF_analysis/seurat_objects/merged.CSF_integrated_harmony_corrected_cluster.rds")

pca <- seurat@reductions$harmony@cell.embeddings

umap <- seurat@reductions$umap@cell.embeddings


meta <- seurat@meta.data



singleR_res_jess <- readRDS("./results/CSF_analysis/reductions/allcells_SingleR_SingleCellScore.rds")

```



## SingleR results per cell and clustering

\  

```{r, fig.width=10}

Mo_barcodes <- meta %>%
  tibble::rownames_to_column("index") %>%
  dplyr::filter(cluster_label == "Mø") %>%
  pull(index)


singleR_Mo <- singleR_res_jess[Mo_barcodes,]


singleR_Mo <- singleR_Mo / matrixStats::rowMaxs(singleR_Mo)


hcluster <- singleR_Mo %>% dist() %>% hclust()

annotation_cells <- data.frame(hcluster = as.factor(cutree(hcluster,2)),
                               hcluster5 = as.factor(cutree(hcluster,5))) %>%
  tibble::rownames_to_column("index") %>%
  left_join(meta %>%
              tibble::rownames_to_column("index") %>%
              dplyr::select(index, SingleR.labels, nCount_RNA,nFeature_RNA, LP)
    ) %>%
  tibble::column_to_rownames("index") %>%
  mutate(SingleR.labels = ifelse(grepl("Th|regu|help",SingleR.labels), "CD4",
                                 ifelse(grepl("CD8",SingleR.labels), "CD8",
                                        ifelse(grepl("B cells",SingleR.labels), "B cells",
                                               ifelse(grepl("mono|Mye",SingleR.labels), "myeloid",
                                                      SingleR.labels))))) %>%
  mutate(nCount_RNA = log10(nCount_RNA),
         nFeature_RNA = log10(nFeature_RNA))
  

Heatmap(
  t(singleR_Mo),
  col = c(viridis(1000), "red"),
  show_column_names = F,
  bottom_annotation = HeatmapAnnotation(
    df = annotation_cells,
    col = list(
      nCount_RNA = colorRamp2(
        quantile(annotation_cells$nCount_RNA, c(0.025, 0.975)),
        c("white", "darkgreen")
      ),
      nFeature_RNA = colorRamp2(
        quantile(annotation_cells$nFeature_RNA, c(0.025, 0.975)),
        c("white", "darkorange2")
      ),
      LP = c(
        Healthy = "grey",
        Baseline = "red",
        Follow_up = "royalblue"
      )
    )
  )
)

```

We can see that there is 2 lineages present in this cluster based on the singleR scores. And we can use hierarchical clustering on those scores to separate the 2 lineages.

## Count and Feature on the clustered populations

\  


```{r}


annotation_cells %>% 
  gather("metric","value", -!matches("_RNA$")) %>%
  ggplot(aes(hcluster, value)) +
  geom_violin() +
  facet_wrap(.~metric, scales = "free", strip.position = "left") +
  theme_bw() +
  theme(strip.placement = "outside",
        axis.title.y = element_blank(),
        strip.background = element_blank()) +
  labs(caption = "log10 scale")

annotation_cells %>% 
  gather("metric","value", -!matches("_RNA$")) %>%
  ggplot(aes(hcluster, 10^value)) +
  geom_violin() +
  facet_wrap(.~metric, scales = "free", strip.position = "left") +
  theme_bw() +
  theme(strip.placement = "outside",
        axis.title.y = element_blank(),
        strip.background = element_blank()) +
  labs(caption = "linear scale")




```

As expected myeloid cells are more transcriptomically active and have larger library sizes on average.

## look at euclidean distances for M-T cluster

\  

If there is truly 2 lineages in this cluster, each lineage should be closer to other cells from the same lineages by euclidean distance in PCA space.

```{r}

hcluster <- as.factor(cutree(hcluster, 2))

meta_sampled <- meta %>%
  tibble::rownames_to_column("index") %>%
  left_join(data.frame(hcluster = hcluster) %>% tibble::rownames_to_column("index")) %>%
  mutate(cluster_label = ifelse(cluster_label == "Mø", paste0(cluster_label, "_", hcluster), as.character(cluster_label))) %>%
  dplyr::filter(grepl("Mø", cluster_label)) %>%
  group_by(cluster_label, LP) %>%
  summarize(count = n())

meta_sampled <- meta %>%
  tibble::rownames_to_column("index") %>%
  left_join(data.frame(hcluster = hcluster) %>% tibble::rownames_to_column("index")) %>%
  mutate(cluster_label = ifelse(cluster_label == "Mø", paste0(cluster_label, "_", hcluster), as.character(cluster_label))) %>%
  group_by(Emulsion, cluster_label) %>%
  sample_n(100, replace = T) %>%
  ungroup() %>%
  distinct()

pca_sampled <- pca[meta_sampled$index,]

distances <- dist(pca_sampled) %>% broom::tidy()

distances <- distances %>%
  left_join(meta_sampled %>% dplyr::select(item1 = index, cluster_label)) %>%
  left_join(meta_sampled %>% dplyr::select(item2 = index, cluster_label_2 = cluster_label))

distances_meta <- distances %>%
  dplyr::filter(cluster_label == "Mø_1" | cluster_label_2 == "Mø_1") 

distances_meta <- distances_meta %>%
  unite("cluster", cluster_label, cluster_label_2) %>%
  mutate(cluster = gsub("Mø_1", "", cluster))


distances_meta <- distances_meta %>%
  mutate(cluster = gsub("^_", "", cluster)) %>%
  mutate(cluster = gsub("_$", "", cluster)) %>%
  mutate(cluster = ifelse(cluster == "", "Mø_1", cluster))


table(distances_meta$cluster)



ggplot(distances_meta, aes(cluster, distance, fill = cluster == "Mø_1")) +
  geom_violin(draw_quantiles = c(0.25,0.5,0.75)) +
  theme_mrl() +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("grey","orange")) +
  ggtitle("euclidean distances to cells in Mø_1 cluster, aka probable T cells")




```


```{r}



distances_meta <- distances %>%
  dplyr::filter(cluster_label == "Mø_2" | cluster_label_2 == "Mø_2") 

distances_meta <- distances_meta %>%
  unite("cluster", cluster_label, cluster_label_2, remove = F) %>%
  mutate(cluster = gsub("Mø_2", "", cluster))


distances_meta <- distances_meta %>%
  mutate(cluster = gsub("^_", "", cluster)) %>%
  mutate(cluster = gsub("_$", "", cluster)) %>%
  mutate(cluster = ifelse(cluster == "", "Mø_2", cluster))


table(distances_meta$cluster)

rm(distances)

ggplot(distances_meta, aes(cluster, distance, fill = cluster == "Mø_2")) +
  geom_violin(draw_quantiles = c(0.25,0.5,0.75)) +
  theme_mrl() +
  theme(legend.position = "none") +
  scale_fill_manual(values = c("grey","orange")) +
  ggtitle("euclidean distances to cells in Mø_2 cluster, aka true myeloid cells")
```
We can see that overall, as expected the T cells are closer to other T cell clusters, and similarly the macrophages are closer to other myeloid subsets. We'll thus rely on the present hierarchical clustering of the singleR scores to separate those 2 lineages and define 2 new clusters, based on the split of the dendrogram into 2 clusters.


```{r, eval=T}


# save seurat object with hierarchical clustering assignments

meta <- meta %>%
  tibble::rownames_to_column("index") %>%
  left_join(data.frame(hcluster = hcluster) %>% tibble::rownames_to_column("index")) %>%
  tibble::column_to_rownames("index")


seurat@meta.data <- meta[colnames(seurat),]

saveRDS(seurat, "results/CSF_analysis/seurat_objects/merged.CSF_integrated_harmony_corrected_cluster_hcluster.rds")
```

