---
title: "New analysis with 2021 samples"
author: "PPA"
date: "`r Sys.Date()`"
output:
  pdf_document: 
    number_sections: yes
    toc: yes
    toc_depth: 4
  html_document:
    df_print: paged
geometry: margin=0.5in
header-includes:
   - \usepackage{subfig}
   - \usepackage{float}
---


```{r setup, include=FALSE}

##########################
## Packages and options ##
##########################

library(gridExtra)
library(knitr)
library(readr)
library(dplyr)
library(tidyr)
library(tibble)
library(stringr)
library(ggplot2)
library(ggExtra)
library(stats)
# library(RColorBrewer)
library(viridis)
library(scales)
library(ggrepel)
library(circlize)
library(RColorBrewer)
library(gplots)
library(cowplot)
library(gtable)
library(data.table)
library(Seurat)
library(hdf5r)
library(ggsci)
library(ggridges)
library(forcats)
library(kableExtra)
library(sp)
# library(FactoMineR)
# library(factoextra)
library(wesanderson)
library(magrittr)
library(ComplexHeatmap)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.pos = 'H')
knitr::opts_chunk$set(dev = 'png', dpi = 300)

#set root directory to be one level up of the current directory
#knitr::opts_knit$set(root.dir = '.')

#Caching option when tweaking the knitting
knitr::opts_chunk$set(cache = T)

WRITE_FILES <- FALSE


##################
## ggplot theme ##
##################

theme_mrl <- function(x = 1) {
  theme_minimal() +
    theme(
      axis.line = element_line(),
      axis.ticks.x = element_line(),
      axis.ticks.y = element_line(),
      axis.text.x = element_text(size = 12*x,face = "bold", angle = 45, vjust = 0.9, hjust=0.9),
      axis.text.y = element_text(size = 12*x,face = "bold"),
      axis.title.x = element_text(size = 12*x,face = "bold"),
      axis.title.y = element_text(size = 12*x,face = "bold"),
      strip.background = element_rect(fill="gray20", colour="gray20", linetype="solid"),
                                      strip.text = element_text(size=12*x, colour="white", face="bold"),
                                      legend.title = element_text(size=14*x, face = "bold"),
                                      legend.text = element_text(size=12*x, color="gray20", face="bold"),
      legend.background = element_rect(fill = "transparent", colour = "transparent"),
      plot.title =  element_text(hjust=0.5, vjust=2, face="bold"),
      plot.subtitle = element_text(hjust=0.5, vjust=3, face="italic"),
      plot.caption = element_text(hjust = 0, face = "italic")
                                      )
}


opts <- options()  # save old options

options(ggplot2.continuous.colour="viridis")
options(ggplot2.continuous.fill = "viridis")

palette <- c(pal_rickandmorty()(12),pal_jco()(10), pal_futurama()(11), pal_tron()(7))


options(bitmapType = 'cairo', device = 'png')
```



```{r, fig.width=10}  
# function to overlay values onto the UMAP embeddings
hex_plots <- function(data,dim1,dim2, color, facets = "", CI = 0.05) {

  # calculate the quantilles for CI limits onthe color scale.
  if(CI) {
    lims <- quantile(data[,color], probs = c(CI/2,1 - CI/2))
  } else {
    lims <- c(NA,NA)
  }
  
  
  ggplot(data, aes_string(dim1, dim2)) +
  stat_summary_hex(bins = 100,
                   fun = median,
                   aes(z= !!ensym(color), color = ..value..)) +
    facet_wrap(as.formula(paste(". ~", facets)), scales = "free") +
    ggtitle(color) +
    theme_mrl() +
    guides(fill = guide_none()) +
    scale_fill_viridis(limits = lims, oob = squish) +
    scale_color_viridis(limits = lims, oob = squish)
}



ridges_plots <- function(data,dim1,dim2) {
  
  ggplot(data, 
       aes_string(dim1, 
                  dim2)) +
  stat_density_ridges(aes(fill = factor(stat(quantile))),
    geom = "density_ridges_gradient", calc_ecdf = TRUE,
    quantiles = c(0.01,0.05, 0.1,0.9,0.95,0.99), quantile_lines = TRUE,
    color = NA) +
  facet_grid(curated_types~., scales = "free_y", space = "free_y") +
  scale_fill_manual(labels = c("0.01","0.05", "0.1","0.1-0.9","0.9","0.95","0.99"),
                    values = c("black","grey80","black","grey80","black","grey80","black"))
}





```




# filter cells QC


```{r, fig.asp= 1.1}




high_point <- function(x, high_only = F) {
  den <- density(x)
  high <- data.frame(x = den[["x"]],
                     y = den[["y"]]) %>%
    dplyr::filter(y == max(y)) %>%
    pull(x) %>%
    mean()
  
  if(!high_only) {
    if(mean(x) < high) {
      return(c(y = high,ymin = mean(x), ymax = high))
    } else {
      return(c(y = high,ymin = high, ymax = mean(x)))
    }
  } else {
    return(high)
  }
  
    
  
}


raw_meta <- readRDS("results/PBMC/seurat_QC/master_meta.rds") %>%
  tibble::rownames_to_column("cell.id") %>%
  separate(cell.id, c("patientID", "t" , "barcode"), sep = "_") %>%
  mutate(t = ifelse(grepl("6m|6M", t), "t6M", "t0M")) %>%
  mutate(sample = paste(patientID, t, sep = "_"))

unique(raw_meta$sample)


raw_meta %>%
  ggplot(aes(nCount_RNA, sample, fill = patientID)) +
  geom_density_ridges(quantile_lines = TRUE) +
  scale_x_log10(limits = c(NA,NA)) +
  stat_summary(fun.data = "high_point", geom = "pointrange", position = position_nudge(y = 0.8)) +
  scale_fill_cyclical(values = c("tan", "lightskyblue3")) +
  annotation_logticks(sides = "b") +
  geom_vline(xintercept = 1200)

raw_meta %>%
  dplyr::filter(nCount_RNA > 1200 & nFeature_RNA > 500 & nFeature_RNA < 4000) %>%
  ggplot(aes(nFeature_RNA, sample, fill = patientID)) +
  geom_density_ridges(quantile_lines = TRUE) +
  stat_summary(fun.data = "high_point", geom = "pointrange", position = position_nudge(y = 0.8)) +
  scale_x_log10() +
  scale_fill_cyclical(values = c("tan", "lightskyblue3")) +
  annotation_logticks(sides = "b")





# center with high point


test <- raw_meta %>%
  dplyr::filter(nCount_RNA > 1200 & nFeature_RNA > 500 & nFeature_RNA < 4000) %>%
  mutate(nFeature_RNA = log10(nFeature_RNA)) %>%
  group_by(sample) %>%
  nest() %>%
  mutate(high = purrr::map(data, ~rep(high_point(.$nFeature_RNA, high_only = T),
                                      nrow(.)))) %>%
  unnest() %>%
  ungroup() %>%
  mutate(nFeature_RNA = 10^(nFeature_RNA - high + mean(nFeature_RNA)))

test %>%
  ggplot(aes(nFeature_RNA, sample, fill = patientID)) +
  geom_density_ridges(quantile_lines = TRUE) +
  stat_summary(fun.data = "high_point", geom = "pointrange", position = position_nudge(y = 0.8)) +
  scale_x_log10() +
  scale_fill_cyclical(values = c("tan", "lightskyblue3")) +
  annotation_logticks(sides = "b") +
  geom_vline(xintercept = c(900, 1800))







```








```{r import rds from Jess}


meta_table <- readRDS("results/PBMC/clustering/master_seurat_curated_types_table_harmony_iterative_batch_emulsion.rds") %>%
  mutate(treatment = ifelse(grepl("6", treatment), "t6M","t0M"),
         curated_clusters = unclass(as.factor(curated_clusters)))



meta_table <- readRDS("results/PBMC/clustering/master_seurat_coarse_types_table_harmony_iterative_batch_emulsion.rds") %>%
  dplyr::select(matches("coarse|cell\\.id")) %>%
  right_join(meta_table)
  

```






\pagebreak



# Low dimensionality embeddings


## Effect of knn, min distance and spread

```{r choosing UMAP parameters, fig.width=10}

initial_meta <- readRDS("results/PBMC/clustering/master_seurat_low_res_clustering_table_harmony_iterative_batch_emulsion.rds") %>%
  mutate(treatment = ifelse(grepl("6", treatment), "t6M","t0M"))

UMAP <- initial_meta %>%
  dplyr::select(matches("^UMAP|cell\\.id")) %>%
  gather("embed", "coord", -cell.id) %>%
  tidyr::separate(embed, c("iteration", "dim"), sep = "_") %>%
  mutate(dim = paste0("dim_", dim)) %>%
  spread(dim, coord) %>%
  separate(iteration, c("knn", "min.dist", "spread"), sep = "x") %>%
  mutate(min.dist = gsub("Md0", "Md0.", min.dist)) %>%
  mutate(min.dist = fct_reorder(min.dist,
                                parse_number(min.dist))) %>%
  mutate(spread = fct_reorder(spread,
                              as.numeric(gsub("S", "", spread))))


for(knn in c(50)) {
  filter = paste0("UMAPn",knn)
  
  plot <- UMAP %>%
    dplyr::filter(knn == filter) %>%
    #mutate(dim_2 = ifelse(knn == "UMAPn50", -dim_2, dim_2)) %>% # invert dim 2 for knn = 50 as it seems flipped
    ggplot(aes(dim_1,
               dim_2)) +
    geom_hex(aes(color = ..count..), bins = 200) +
    scale_color_viridis(trans = "log2") +
    scale_fill_viridis(trans = "log2") +
    facet_wrap(min.dist~spread, scales = "free") +
    theme(axis.line = element_blank(),
          axis.ticks = element_blank(),
          axis.title = element_blank(),
          axis.text = element_blank()) +
    ggtitle(paste("knn =",knn))
  
  cat("\n\n")
  
  print(plot)
  
  cat("\n\n")
  
}



```

\pagebreak



## Clusters at low resolution


```{r low res clusters, fig.width=10, fig.asp= 1}



best_clusters <- initial_meta %>%
  dplyr::select(cell.id, SNNk20_algo1_res0.1)


for (knn in c(50)) {
  filter = paste0("UMAPn", knn)
  
  
  plot <- UMAP %>%
    dplyr::filter(knn == filter & grepl("5", spread)) %>%
    left_join(best_clusters) %>%
    #mutate(dim_2 = ifelse(knn == "UMAPn50", dim_2, -dim_2)) %>% # invert dim 2 for knn = 50 as it seems flipped
    ggplot(aes(dim_1,
               dim_2)) +
    geom_hex(aes(fill = SNNk20_algo1_res0.1, alpha = log(..ndensity..)),
             bins = 200,
             lwd = 0.01) +
    scale_alpha(range = c(0.2, 1)) +
    theme_mrl() +
    facet_wrap(min.dist~spread, scales = "free", ncol = 2) +
    theme(panel.background = element_rect(fill = 'grey90', color = NA),
          legend.position = "bottom") +
    scale_fill_manual(values = palette) +
    scale_color_manual(values = palette) +
    ggtitle(knn)
  
  cat("\n\n")
  
  print(plot)
  
  cat("\n\n")
  
}

```

\pagebreak


## Clusters at intermediate resolution, and comparison with UMAP parameters 

\  

For UMAP I fixed: knn = 50 and spread = 5, looking at variation on min distance


```{r higher res clusters, fig.width=10, fig.asp= 1}

best_clusters <-initial_meta %>%
  dplyr::select(cell.id, SNNk20_algo1_res0.2) %>%
  mutate(SNNk20_algo1_res0.2 =as.factor(SNNk20_algo1_res0.2))

for (knn in c(50)) {
  filter = paste0("UMAPn", knn)
  
  
  plot <- UMAP %>%
    dplyr::filter(knn == filter & grepl("5", spread)) %>%
    left_join(best_clusters) %>%
    #mutate(dim_2 = ifelse(knn == "UMAPn50", dim_2, -dim_2)) %>% # invert dim 2 for knn = 50 as it seems flipped
    ggplot(aes(dim_1,
               dim_2)) +
    geom_hex(aes(fill = as.factor(SNNk20_algo1_res0.2), alpha = log(..ndensity..)),
             bins = 200,
             lwd = 0.01) +
    scale_alpha(range = c(0.2, 1)) +
    theme_mrl() +
    facet_wrap(min.dist~spread, scales = "free", ncol = 2) +
    theme(panel.background = element_rect(fill = 'grey90', color = NA),
          legend.position = "bottom") +
    scale_fill_manual(values = palette) +
    scale_color_manual(values = palette) +
    ggtitle(knn)
  
  cat("\n\n")
  
  print(plot)
  
  cat("\n\n")
  
}


```


\pagebreak


## Clusters at higher resolution, and comparison with UMAP parameters 

\  

For UMAP I fixed: knn = 50 and spread = 5, looking at variation on min distance

this resolution seems too high

```{r higher res clusters, fig.width=10, fig.asp= 1}

best_clusters <-meta_table %>%
  dplyr::select(cell.id, SNNk20_algo1_res0.5)

for (knn in c(50)) {
  filter = paste0("UMAPn", knn)
  
  
  plot <- UMAP %>%
    dplyr::filter(knn == filter & grepl("5", spread)) %>%
    left_join(best_clusters) %>%
    #mutate(dim_2 = ifelse(knn == "UMAPn50", dim_2, -dim_2)) %>% # invert dim 2 for knn = 50 as it seems flipped
    ggplot(aes(dim_1,
               dim_2)) +
    geom_hex(aes(fill = SNNk20_algo1_res0.5, alpha = log(..ndensity..)),
             bins = 200,
             lwd = 0.01) +
    scale_alpha(range = c(0.2, 1)) +
    theme_mrl() +
    facet_wrap(min.dist~spread, scales = "free", ncol = 2) +
    theme(panel.background = element_rect(fill = 'grey90', color = NA),
          legend.position = "bottom") +
    scale_fill_manual(values = palette) +
    scale_color_manual(values = palette) +
    ggtitle(knn)
  
  cat("\n\n")
  
  print(plot)
  
  cat("\n\n")
  
}


```


\pagebreak





## QC inspection


```{r looking at cluster 9, fig.asp=1, fig.width=15}

QC_metrics <- initial_meta %>%
  dplyr::select(matches("SNNk20_algo1_res0\\.2|cell.id|_RNA|donor|treatment")) %>%
  mutate(SNNk20_algo1_res0.2 =as.factor(SNNk20_algo1_res0.2))


best_UMAP <- UMAP %>%
    dplyr::filter(knn == "UMAPn50" & 
                    #grepl("5",spread) & 
                    min.dist ==  "Md0.01" & 
                    spread == "S5")

# save for later
saveRDS(best_UMAP, "results/PBMC/clustering/best_UMAP.rds")


best_clusters <-meta_table %>%
  dplyr::select(cell.id, SNNk20_algo1_res0.2)


cat("\n\n")

p1 <- ggplot(QC_metrics, 
       aes(SNNk20_algo1_res0.2, nCount_RNA)) +
  geom_boxplot() +
  annotation_logticks(sides = "l") +
  scale_y_log10()

cat("\n\n")

p2 <- ggplot(QC_metrics, 
       aes(SNNk20_algo1_res0.2, nFeature_RNA)) +
  geom_boxplot() +
  annotation_logticks(sides = "l") +
  scale_y_log10()




p3 <- best_UMAP %>%
  #left_join(best_clusters) %>%
  #dplyr::filter(SNNk20_algo1_res0.3 == 9) %>%
  left_join(QC_metrics) %>%
  ggplot(aes(dim_1,
               dim_2)) +
  # stat_bin_hex(data = best_UMAP, 
  #                color = "grey",
  #                fill = "grey",
  #                bins = 100) +
  #facet_wrap(.~SNNk20_algo1_res0.3, ncol = 3) +
  stat_summary_hex(aes(z = nFeature_RNA), bins = 100) +
  scale_fill_viridis(trans = "log10", limits = c(500,3800)) +
  guides(fill = guide_colorbar(title = "nFeature_RNA")) +
  ggtitle("Number of Features detected in UMAP embeddings")


cat("\n\n")

plot_grid(p1,p2,p3,
          ncol = 1,
          rel_heights = c(0.5,0.5,1))


cat("\n\n")

QC_metrics %>%
  ggplot(aes(paste(donor,treatment),
             fill = donor,
             color = treatment)) +
  geom_bar(alpha = 0.7) +
  theme_mrl() +
  scale_fill_manual(values = palette) +
  scale_color_manual(values = c(NA, "black")) +
  theme(legend.position = "bottom",
        axis.text.x = element_blank()) +
  facet_wrap(.~SNNk20_algo1_res0.2, ncol = 3, scales = "free_y") +
  labs(title = "Cell # contributions per emulsion",
       x = "emulsion",
       y = "cell #")

```



# Reference calls and subclustering based on singleR reference calls

## Running singleR on clusters

```{r, fig.width=15}




singleR_res <- readRDS("results/PBMC/clustering/subclustered_master_seurat_harmony_iterative_batch_emulsion_singleR.rds")


for(singleR_output in names(singleR_res)) {
  
  
  
  title <- singleR_output
  
  print(title)
  
  singleR_output <- singleR_res[[singleR_output]]

  rownames(singleR_output[[1]][[1]]) <- rownames(singleR_output[[1]][[2]])
  
  
  monaco.matrix <- singleR_output[[1]][[1]] / matrixStats::rowMaxs(singleR_output[[1]][[1]])
  
  monaco.labels <- (monaco.matrix == 1) * singleR_output[[1]][[1]] %>% round(digits = 2)
  
  monaco.labels[monaco.labels == 0] <- ""
  
  new_cluster_names <- paste(unclass(as.factor(rownames(monaco.matrix))), rownames(monaco.matrix), sep = "   ")
  
  rownames(monaco.matrix) <- new_cluster_names
  
  pheatmap::pheatmap(t(monaco.matrix), 
                     scale = "none", 
                     border_color = NA,
                     color = magma(100),
                     cutree_cols = 4,
                     cutree_rows = 4,
                     display_numbers = t(monaco.labels),
                     main = title)
  
  
  
  cat("\n\n")
  
  rownames(singleR_output[[2]][[1]]) <- rownames(singleR_output[[2]][[2]])
  
  dice.matrix <- singleR_output[[2]][[1]] / matrixStats::rowMaxs(singleR_output[[2]][[1]])
  
  dice.labels <- (dice.matrix == 1) * singleR_output[[2]][[1]] %>% round(digits = 2)
  
  dice.labels[dice.labels == 0] <- ""
  
  new_cluster_names <- paste(unclass(as.factor(rownames(dice.matrix))), rownames(dice.matrix), sep = "   ")
  
  rownames(dice.matrix) <- new_cluster_names
  
  pheatmap::pheatmap(t(dice.matrix), 
                     scale = "none", 
                     border_color = NA,
                     color = magma(100),
                     cutree_cols = 4,
                     cutree_rows = 4,
                     display_numbers = t(dice.labels),
                     main = title)
}




```

`meta_subC` is the manual curation, choosing different resolutions depending on the subclustering groups. I then renumber those calls from 1 to 38 in `curated_clusters`

```{r higher res clusters, fig.width=10, fig.asp= 1}


best_clusters <- meta_table %>%
  dplyr::select(cell.id, curated_clusters) %>%
  mutate(curated_clusters= as.factor(curated_clusters))


plot <- best_UMAP %>%
  left_join(best_clusters)

cluster_labels <- plot %>%
  group_by(curated_clusters) %>%
  summarize(dim_1 = median(dim_1), dim_2 = median(dim_2))

plot %>%
  ggplot(aes(dim_1,
             dim_2)) +
  geom_hex(aes(fill = curated_clusters, alpha = log(..ndensity..)),
           bins = 300,
           lwd = 0.001) +
  geom_label_repel(data = cluster_labels, 
             aes(label = curated_clusters, fill = as.character(curated_clusters)), 
             color = "black", size =5, alpha = 0.8) +
  scale_alpha(range = c(0.2, 1)) +
  theme_mrl() +
  facet_wrap(min.dist ~ spread, scales = "free", ncol = 2) +
  theme(panel.background = element_rect(fill = 'grey90', color = NA)) +
  scale_fill_manual(values = palette) +
  scale_color_manual(values = palette) +
  ggtitle(knn)


```


## Check singleR results at single cell level for cell type sanity check

Evaluate contamination with singleR reference typing at the single cell level. All cells score high on monocyte references, but some also score high for other lineages. This suggest some hybrid cell type or doublet.


```{r, eval=F, fig.width=10}
#####
mon_sc <- readRDS("results/PBMC/clustering/cell_typing_sc_curated_types_master_seurat_harmony_iterative_batch_emulsion.rds")[[1]]


rownames(mon_sc[[1]]) <- rownames(mon_sc[[2]])

mon_sc <- mon_sc[[1]] / matrixStats::rowMaxs(mon_sc[[1]])



cluster_ids <- meta_table %>%
  mutate(cluster = paste(curated_clusters, fine_clusters, sep = "_")) %>%
  dplyr::select(cluster, cell.id) %>%
  group_by(cluster) %>%
  summarize(ids = list(cell.id)) %>%
  deframe()
    
hm_list <- lapply(names(cluster_ids), function(cluster) {
  cluster_type_scores <- mon_sc[cluster_ids[[cluster]], ]
  
  
  hm <- Heatmap(
    t(cluster_type_scores),
    col = c(viridis(1000), "#ff0000"),
    column_title = cluster,
    show_column_names = F,
    column_km = ceiling(length(cluster_ids[[cluster]])/10000)
  )
  
  png(file.path("results","clustering", paste0(cluster, ".png")))
  
  draw(hm)
  
  dev.off()
  
  return(hm)
})


cluster@meta.data <- cluster_type_scores %>%
  kmeans(2) %>%
  extract2("cluster") %>%
  data.frame(km = .) %>%
  tibble::rownames_to_column("cell.id") %>%
  right_join(cluster@meta.data %>% tibble::rownames_to_column("cell.id")) %>%
  tibble::column_to_rownames("cell.id")

meta_cluster <- cluster@meta.data %>%
  dplyr::select(c("nFeature_RNA","nCount_RNA","t6M","treatment", "NK", "km")) %>%
  mutate(nFeature_RNA = log10(nFeature_RNA), nCount_RNA = log10(nCount_RNA), km = paste0("km",km))




cluster_type_scores <- cluster_type_scores / matrixStats::rowMaxs(cluster_type_scores)

row_clust <- hclust(dist(t(cluster_type_scores)))

col_clust <- hclust(dist(cluster_type_scores))

cluster_split <- paste0("cluster ", cutree(col_clust, k = 2))

names(cluster_split) <- rownames(cluster_type_scores)
cluster_col <- ggsci::pal_futurama()(2)

names(cluster_col) <- unique(cluster_split)





```


# sample representation across clusters




```{r, fig.width=15}


ggplot(meta_table, aes(as.factor(curated_clusters), fill = paste(donor, treatment))) +
  geom_bar() +
  scale_fill_manual(values = palette) +
  guides(fill = guide_legend(ncol = 1)) +
  theme_mrl(0.7)



ggplot(meta_table, aes(as.factor(curated_clusters), fill = paste(donor, treatment))) +
  geom_bar(position = "fill") +
  theme_mrl() +
  scale_fill_manual(values = palette)

```


# gene markers

## against all

```{r, fig.width=10}


presto_results <- readRDS("results/PBMC/DGE/presto_markers_curated_clusters.rds")


presto_results <- do.call(rbind, presto_results) %>%
  dplyr::filter(padj < 0.01)


ggplot(presto_results, aes(group, auc)) +
  geom_violin() +
  geom_boxplot(width = 0.3, 
               outlier.alpha = 0.5,
               outlier.size = 0.5) +
  theme_mrl() +
  geom_hline(yintercept = c(0.65, 0.35), linetype = "dashed")


# number of markers kept
presto_results %>% 
  dplyr::filter(auc < 0.35 | auc > 0.65) %>%
  pull(feature) %>%
  unique() %>%
  length()

top_markers <- presto_results %>% 
  dplyr::filter(auc < 0.35 | auc > 0.65) %>%
  group_by(group) %>%
  mutate(up = ifelse(auc > 0.5,rank(-auc), NA)) %>%
  mutate(dn = ifelse(auc < 0.5,rank(auc), NA)) %>%
  ungroup() %>%
  dplyr::filter(up < 11 | dn < 11) %>%
  group_by(group) %>%
  nest() %>%
  ungroup() %>%
  mutate(up = purrr::map(data, ~ { .x %>%
      dplyr::filter(up < 11) %>%
      arrange(-auc) %>%
      pull(feature) %>%
      paste(collapse = ", ")  
  })) %>%
  mutate(dn = purrr::map(data, ~ { .x %>%
      dplyr::filter(dn < 11) %>%
      arrange(auc) %>%
      pull(feature) %>%
      paste(collapse = ", ")  
  })) %>%
  dplyr::select(-data) %>%
  unnest() %>%
  mutate(cluster = unclass(as.factor(group)))
  

plot_grid(grid.table(top_markers))

```




## paired markers

```{r, eval=F}
curation <- readRDS("results/PBMC/clustering/master_seurat_curated_types_table_harmony_iterative_batch_emulsion.rds") %>%
  mutate(curated_clusters = unclass(as.factor(meta_subC))) %>%
  dplyr::select(curated_types, curated_clusters, meta_subC) %>%
  distinct()


paired_presto <- readRDS("results/PBMC/DGE/presto_paired_markers_curated_clusters.rds")


paired_presto <- do.call(rbind, paired_presto) %>%
  dplyr::filter(padj < 0.05)


saveRDS(paired_presto, "results/PBMC/DGE/presto_paired_markers_curated_clusters_filtered.rds")
```


```{r, fig.width=10}


paired_presto <- readRDS("results/PBMC/DGE/presto_paired_markers_curated_clusters_filtered.rds")


paired_presto %>%
  dplyr::filter(auc < 0.4 | auc > 0.6) %>%
  dplyr::select(feature) %>%
  distinct() %>%
  pull(feature) %>%
  length()

paired_presto %>%
  mutate(pctFC = log2((pct_in + 0.01) / (pct_out + 0.01))) %>%
  dplyr::select(pctFC) %>%
  ggplot(aes(pctFC)) +
  geom_histogram(bins = 100)





lib_sizes <- readRDS("results/PBMC/clustering/master_seurat_curated_types_table_harmony_iterative_batch_emulsion.rds") %>%
  mutate(curated_clusters = unclass(as.factor(meta_subC))) %>%
  dplyr::select(curated_types, curated_clusters, meta_subC, cell.id, nFeature_RNA, nCount_RNA, sample) %>%
  group_by(sample) %>%
  mutate(nFeat_quantiles = ntile(nFeature_RNA, 100))
  

ggplot(lib_sizes, aes(meta_subC,as.numeric(nFeat_quantiles), fill = sample)) +
  stat_summary_bin(position = "jitter") +
  theme_mrl() +
  facet_grid(.~curated_types, scales = "free_x", space = "free_x")

```



